---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# augury

<!-- badges: start -->
[![R-CMD-check](https://github.com/caldwellst/augury/workflows/R-CMD-check/badge.svg)](https://github.com/caldwellst/augury/actions)
<!-- badges: end -->

The goal of augury is to streamline the process of fitting models to infill and
forecast data, particularly for models used within the WHO's Triple Billion
framework

## Installation

You can install augury from [GitHub](https://github.com/caldwellst/augury). The
package depends on the [INLA package](https://www.r-inla.org/home) which is not
available on CRAN. You will need to separately install that prior to installing
augury. The below code should allow you to install the packages without any
trouble:

``` r
if (!require("INLA")) install.packages("INLA",repos=c(getOption("repos"),INLA="https://inla.r-inla-download.org/R/stable"), dep=TRUE)
remotes::install_github("caldwellst/augury")
```

If you are reinstalling the package to update something, you should make sure you
have necessary dependencies, but to avoid re-downloading the large INLA tarball,
you can use:
```r
remotes::install_github("caldwellst/augury", dependencies = FALSE)
```

## Overview

Most of the functions available through the package are designed to streamline
fitting a model and replacing missing observations within a single data frame
in one pass.

* `predict_general_mdl()` is a function that takes in a data frame, generic R
    modelling function (such as `stats::lm`), and fits a model and returns newly
    projected data, error metrics, the fitted model itself, or a list of all 3.
* `predict_inla()` is a function similar to the above, except rather than taking
    a generic modelling function it accepts a formula and additional argument
    passed to `INLA::inla()` to perform Bayesian analysis of structured additive
    models.
* `predict_forecast()` instead uses forecasting methods available from
    `forecast::forecast()` to perform exponential smoothing or other time series
    models on a response variable.
    
Given that models might need to be fit separately to different portions of a data
frame, such as fitting time series models for each country individually, these
functions now accept the argument `group_models` that determines whether to fit
separate models to each group in the data frame before joining them back together
into the original data frame.

## Additional model wrappers

Additional functions are provided as requested to streamlining fitting of
specific models. For general modeling, there are grouped and individual functions
for `stats::lm()`, `stats::glm()`, and `lme4::lmer()` to fit linear models,
generalized linear models, and linear mixed-effects models respectively. These
wrappers are:

* `predict_lm()`
* `predict_glm()`
* `predict_lmer()

As well, wrappers are provided around INLA for models currently in place for the
Triple Billion framework. These are a time series model with no additional
covariates (fit individually by country) and a mixed-effects model using
covariates.

* `predict_inla_me()`
* `predict_inla_ts()`

And for forecasting, generic functions are provided for simple exponential smoothing
and Holt's linear trend exponential smoothing.

* `predict_holt()`
* `predict_ses()`

## Covariates

For convenience, the covariates used within the default INLA modeling are exported
from the package and can be easily joined up to a data frame for use in modeling.
These are available through `augury::covariates_df`.

## Additional functionality

To help streamline other portions of the modeling process, some other functions
are included in the package.

* `probit_transform()` uses a probit transformation on select columns of a data
   frame, and inverses it if specified.
* `scale_transform()` scales a vector by a single number, very simple, and can
   inverse the scaling if specified.
* `predict_simple()` performs linear interpolation or flat extrapolation in the
   same manner as the other `predict_...` functions, but without modeling or
   confidence bounds.
* `predict_average()` performs averaging by groups of columns in the same manner
   as the other `predict_...` functions, but without modeling or confidence
   bounds.
   
Together, they can be used to transform and scale data for better modeling, and
then inverse these to get data back into the original feature space.

## INLA modeling examples
 
```{r child = "vignettes/inla-modeling.Rmd"}
```

## Forecasting examples

```{r child = "vignettes/forecast-modeling.Rmd"}
```

## Simple prediction methods

```{r child = "vignettes/simple-methods.Rmd"}
```
