---
title: "Average trend modeling"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Average trend modeling}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

While we most often want to directly build models on our original dataset to generate predicted values, we might instead want to generate average trends across larger groups instead, and then apply this to our original data. For instance, generating trends by region, and then applying those regional trends back to the country level. The `predict_...avg_trend()` functions in augury allow us to do just that, applying any of the models we are used to a grouped set of columns.

These work across specific groups, specified by `average_cols`, and averaging numeric values specified as the response variable or variables extracted from a `formula`. The specified model is then fit to this averaged data, and the predicted values are joined back up to the original data frame. Let's look at an example using blood pressure data, which has a comprehensive time series.

```{r inla-avg-setup}
library(augury)

df <- ghost::gho_data("BP_04", query = "$filter=Dim1 eq 'MLE' and Dim2 eq 'YEARS18-PLUS'") %>%
  billionaiRe::wrangle_gho_data() %>%
  dplyr::right_join(covariates_df) %>%
  dplyr::select(iso3, year, year_n, value) %>%
  dplyr::filter(whoville::is_who_member(iso3),                # keep WHO member states
                year >= 2000, year <= 2023) %>%               # get relevant years  
  dplyr::mutate(who_region = whoville::iso3_to_regions(iso3)) # get WHO regions

ur <- unique(df$who_region)
ur
```

Alright, so, here we have `r length(ur)` WHO regions. We will use these regions to fit a model to and use INLA to predict out to 2023, then apply these trends to input countries.

```{r inla-avg-example}
pred_df <- df %>%
  predict_inla_avg_trend(formula = value ~ f(year_n, model = "rw2"),
                         average_cols = c("who_region", "year_n"),
                         group_models = TRUE,
                         group_col = "iso3",
                         sort_col = "year_n")

pred_df %>%
  dplyr::filter(iso3 == "AFG", year >= 2013)
```
Above, we can see we have a generated a model using 2nd order random walk with INLA, however, the model was generated by averaging data to WHO regions first, fitting the random walk to each reach (since `group_models = TRUE`) and then fitting those trends to the original data. Note some specifics of what had to be set, as the `predict_..._avg_trend()` functions are slightly more complex than others:

* `average_cols` must contain the `sort_col`. So, since we use `year_n` in the time series rather than `year`, we will sort by that this time.
* `average_cols` refers to the groupings used for averaging (we take the average for each WHO region and year in this case). Then, the model is fit to `average_cols` that are NOT the `sort_col`.
* `group_col` is the groupings used on the *original data frame*, which is still necessary here when applying the trend back to the original data.

While slightly more complex, ensuring you follow the above means you should easily and successfully get out meaningfull trend predictions for your data frames using trends generated on grouped data.
